---
title: "Model fitting and producing short-term forecasts using daily data with EpiEstim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model fitting and producing short-term forecasts using daily data with EpiEstim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(dplyr)
library(ViroReportR)
library(ggplot2)
library(purrr)
library(tidyr)
```

## Overview 

This vignette provides a short demonstration of the model fitting functions in `ViroReportR` which extend the functionality and wrap the `estimate_R` function from the `EpiEstim` package to produce short-term forecasts. 

## Loading data

For this short demonstration, we will simulate daily data for Influenza A, RSV and Covid-19, using the function `simulate_data` which is provided with the `ViroReportR` package. The input dataset must have two columns: `date` and `confirm` in accordance to format accepted by the model fitting functions.

```{r }
data <- simulate_data(days=365, #days spanning simulation
                      peaks = c("flu_a"=90,"rsv"=110,"sars_cov2"=160), #peak day for each disease
                      amplitudes=c("flu_a"=50,"rsv"=40,"sars_cov2"=20), #amplitude of peak for each disease
                      scales = c("flu_a"=-0.004,"rsv"=-0.005,"sars_cov2"=-0.001), # spread of peak for each disease
                      time_offset = 0, #number of days to offset start of simulation
                      noise_sd = 5, #noise term
                      start_date = "2024-01-07" #starting day (Sunday)
                      )
```

```{r}
data$date <- lubridate::ymd(data$date)

vri_data <- data %>% 
            pivot_longer(cols = -date,  # all columns except 'date'
                          names_to = "disease_type",
                          values_to = "confirm") # <-- need ot be called confirm

# VRI data set-up
vri_name_list <- vri_data %>% 
    dplyr::group_by(disease_type) %>% 
  dplyr::group_keys() %>% pull()


vri_data_list <- vri_data %>% 
  dplyr::group_by(disease_type) %>% 
  dplyr::group_map(~.x)

names(vri_data_list) <- vri_name_list

```

## Model fitting and forecasting

The code below estimates the reproduction number using `EpiEstim` through the `generate_forecast()` function for each disease type via the `purrr::map2()` function. The `generate_forecast()` function prepares the data, estimates the reproduction number, and produces short-term forecasts of daily confirmed cases for an `n_days` forecast horizon.

```{r}
#parameters set-up
start_date <- min(vri_data$date) + 13
forecast_horizon <- 7
smooth <- FALSE
validate_window_size <- 7

forecasts_results <- tibble(
  vri_data_list,
  forecasts = map2(
    vri_data_list,
    vri_name_list,
    ~ generate_forecast(
      data = .x,
      smooth_data = smooth,
      type = .y,
      n_days = forecast_horizon,
      start_date = start_date
    )
  )
)

names(forecasts_results$forecasts) <- vri_name_list
names(forecasts_results$vri_data_list) <- vri_name_list
```

The function supports forecasting for SARS-CoV2, RSV and Influenza A primarily but also provides functionality to forecast other respiratory diseases. Any other respiratory disease can be specified by setting `type = other`. If `type = other` is set, a user-specified `mean_si` and `std_si` must be passed to the function as well. 

## Validation

The package provides validation functionality to support predictive performance evaluation. The forecasting model is trained on historical data and used to predict daily cases r `forecast_horizon` days ahead. Validation compares the model’s forecasted values with observed historical data, where overlap between prediction intervals and actual cases indicates good model performance. A rolling window approach is applied to the historical data to generate multiple validation datasets.



```{r validation-plots, echo=FALSE, results='asis'}
validation_results <- tibble(
  vri_data_list,
  validation = map2(
    vri_data_list,
    vri_name_list,
    ~ generate_validation(
      data = .x,
      type = .y,
      n_days = forecast_horizon,
      start_date = min(.x$date, na.rm = T),
      validate_window_size = validate_window_size,
      window_size = 7,
      smooth_data = smooth,
      smoothing_cutoff = 10
    )
  )
)
names(validation_results$validation) <- vri_name_list
names(validation_results$vri_data_list) <- vri_name_list

for (vri in vri_name_list) {
  cat("\n\n")                               # spacing
  cat(glue::glue("#### {vri}\n\n"))         # subheading for each disease
  print(
    plot_validation(
      data = vri_data_list[[vri]],
      validation_results$validation[[vri]],
      pred_plot = "ribbon"
    )
  )
  cat("\n\n")
}

```

## Forecast Report

Finally, the `ViroReportR` package can conveniently generate an automated report for the current season for all supported viral respiratory diseases (Influenza-A, RSV and SARS-CoV2) using the `generate_forecast_report` function, which renders an HTML report.

```{r eval = FALSE}
tmp_dir <- tempdir()

# Save the simulated data
data_path <- file.path(tmp_dir, "simulated_data.csv")
write.csv(vri_data, data_path, row.names = FALSE)

output_path <- tempdir()
generate_forecast_report(input_dir = data_path,
                            output_dir = output_path,
                            n_days = 7,
                            validate_window_size = 7,
                            smooth = TRUE)

cat("Report saved to:", output_path, "\n")

if (interactive() && file.exists(output_path)) {
  utils::browseURL(output_path)
}

```

