---
title: "Summary report of short-term forecasts for viral respiratory diseases by `ViroReportR`"
output: 
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
params:
  filepath: "O:\\BCCDC\\Groups\\Data_Linked\\Archive\\PANDA\\nb0077\\New_Data/all_disease_data.csv"
  start_date: NULL
  n_days: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
# setup -------------------------------------------------------------------
library(dplyr)
library(ViroReportR)
library(ggplot2)

forecast_horizon <- params$n_days
```

## Overview

This document provides an overview and usage guide for the `ViroReportR` package. The report presents the forecast of `r forecast_horizon`-days ahead confirmed cases for the 3 prevalent respiratory viral cases in BC (Flu-A, RSV and SARS-CoV-2).

`ViroReportR` package uses `EpiEstim` as the backend to forecast daily confirmed cases, which forecasts cases by estimating the instantaneous reproduction number using the [Cori method](https://academic.oup.com/aje/article/178/9/1505/89262) and then simulating the number of new cases using fixed exponential growth over a given time period. This model is designed for short-term forecasting, and its predictive accuracy decreases when extending beyond the optimal 7-14 days horizon.

## Key Functionality

### Core Forecasting Function

#### `forecast_epiestim()`

**Purpose:** Main wrapper function that combines data cleaning, model fitting, and forecast generation into one step.

**Key Steps** 

1. Checks data format and ensures there are at least 14 days of data. 
2. Filters data after the provided `start_date`. 
3. Removes pre-epidemic zero-case days. 
4. Optionally smooths data using `smooth_model_data()`. 
5. Fits reproduction model and generates forecasts. 
6. Summarizes results with quantile estimates.

**Arguments**

| Argument             | Description                                                               |
|--------------------------------|----------------------------------------|
| **data**             | Data frame with `date` and `confirm` columns.                             |
| **start_date**       | Start date of epidemic period.                                            |
| **window_size**      | Rolling window size for $R_t$ estimation.                                 |
| **n_days**           | Forecast horizon (default = 7 days).                                      |
| **type**             | Epidemic type (`"flu_a"`, `"flu_b"`, `"rsv"`, `"sars_cov2"`, `"custom"`). |
| **smooth_data**      | Logical; whether to smooth input curve.                                   |
| **smoothing_cutoff** | Minimum rows required for smoothing (default = 10).                       |
| **...**              | Extra parameters passed to `fit_epiestim_model()`.                        |

**Returns**

Data frame containing: 
  - `date`: forecast date
  - `p025`, `p25`, `p50`, `p75`, `p975`: forecast quantiles
  - `min_sim`, `max_sim`: forecast range

### Supporting Functions

| Function                   | Description                                                                                  |
|---------------------------------|---------------------------------------|
| **`fit_epiestim_model()`** | Function to estimate the reproduction number of epidemics to support short-term forecasts.   |
| **`generate_forecasts()`** | Produce short-term daily projections.                                                        |
| **`smooth_model_data()`**  | Applies P-spline smoothing using a Generalized Additive Model (GAM) to the input model data. |

### Model Validation and Visualization

#### `forecast_time_period()`

**Purpose:** This function runs the forecasting routine across a specified evaluation window. It returns a structured object containing both observed and predicted values for each evaluation period.

**Arguments**

| Argument        | Description                                                              |
|--------------------------------|----------------------------------------|
| **data**        | Data frame with `date` and `confirm` columns.                            |
| **start_date**  | Beginning of the evaluation window.                                      |
| **time_period** | Evaluation granularity ("daily", "weekly", etc.).                        |
| **n_days**      | Forecast horizon (default = 7 days).                                     |
| **type**        | Disease type (`"flu_a"`, `"flu_b"`, `"rsv"`, `"sars_cov2"`, `"custom"`). |
| **algorithm**   | Forecasting algorithm used --- here, "EpiEstim".                         |
| **...**         | Extra parameters passed to `fit_epiestim_model()`.                       |

#### `plot_validation()`

This function visualizes the comparison between observed and predicted incidence.
The pred_plot = "ribbon" option displays prediction intervals as shaded regions around the median forecast.

The resulting plot includes:
  - Black points: observed incidence.
  - Blue line: median predicted incidence (p50).
  - Shaded ribbon: forecast uncertainty (e.g., 25th–75th percentile).

## Example Workflow

### Create simulate data

In this workflow, we simulate RSV incidence data, format it for analysis, and generate short-term forecasts using the EpiEstim-based forecasting functions. We demonstrate the impact of smoothing on forecast results.

```{r}
# create simulate RSV data
disease_type <- "rsv"
test_data <- simulate_data(noise_sd=5)
formatted_data <- get_aggregated_data(
  test_data,
  number_column = disease_type,
  date_column = "date",
  start_date = "2024-04-01",
  end_date = "2024-05-01"
)
```

### Running Forecasts

We generate two forecasts:

With smoothing -- reduces noise in observed case counts to stabilize the {R_t}estimation.

Without smoothing -- uses raw counts directly, showing more variability in predictions.

```{r}
# Run forecasting with smoothing
res_smooth <- forecast_epiestim(
  data = formatted_data,
  start_date = "2024-04-01",
  n_days = forecast_horizon,
  type = "rsv",
  smooth_data = T
)

# Run forecasting without smoothing
res_non_smooth <- forecast_epiestim(
  data = formatted_data,
  start_date = "2024-04-01",
  n_days = forecast_horizon,
  type = "rsv",
  smooth_data = F
)
```

### Viewing Forecast Results

The resulting objects (`res_smooth` and `res_non_smooth`) are lists containing daily forecast quantiles, including: - p025 -- 2.5th percentile - p25 -- 25th percentile - p50 -- median prediction - p75 -- 75th percentile - p975 -- 97.5th percentile

Each row corresponds to a forecasted day, and the columns contain the forecast median and uncertainty intervals.

***Smoothed Forecast Result***
```{r}
res_smooth
```

***Non-Smoothed Forecast Result***
```{r}
res_non_smooth
```

### Visualizing Forecasts

We can visualize forecasts using ribbon plots, where:

-   The shaded ribbon represents the 50% prediction interval.
-   The blue line shows the median forecast.
-   Observed case counts are plotted as black points for comparison.

***Smoothed Forecast Plot***

```{r}
ggplot(res_smooth, aes(x = date)) +
  geom_ribbon(aes(ymin = p25, ymax = p75), fill = "skyblue", alpha = 0.4) +
  geom_line(aes(y = p50), color = "blue", size = 1) +
  geom_point(data = formatted_data, aes( x = date, y = confirm), color = "black", size = 1) +
  labs(
    title = paste0(forecast_horizon, "-Day Forecast of Daily RSV Cases (Smoothed)"),
    x = "Date", y = "Predicted Cases"
  ) +
  theme_minimal()
```

***Non-Smoothed Forecast Plot***

```{r}
ggplot(res_non_smooth, aes(x = date)) +
  geom_ribbon(aes(ymin = p25, ymax = p75), fill = "skyblue", alpha = 0.4) +
  geom_line(aes(y = p50), color = "blue", size = 1) +
  geom_point(data = formatted_data, aes( x = date, y = confirm), color = "black", size = 1) +
  labs(
    title = paste0(forecast_horizon, "-Day Forecast of Daily RSV Cases (Non-Smoothed)"),
    x = "Date", y = "Predicted Cases"
  ) +
  theme_minimal()
```

## Validation

This section demonstrates how to validate the forecast model by comparing predicted values against observed data over a specified time period.

```{r}
time_period_result_daily <- forecast_time_period(
  data = formatted_data,
  start_date = "2024-04-01",
  n_days = forecast_horizon,
  type = "rsv",
  time_period = "daily",
  algorithm = "EpiEstim"
)
```

The validation results can be visualized using plot_validation(), which overlays predicted intervals and observed data points.

```{r}
plot_validation(time_period_result_daily, pred_plot = "ribbon") +
  ggplot2::coord_cartesian(ylim=c(0,90),expand=FALSE)
```

